---
title: "Assignment 01"
author: "Amyre Wells"
date: "2025-01-07"
output:
  pdf_document: default
  html_document: default
---

[Git Repository](https://github.com/amyrewells/biol432_As01)

## Set Up

```{r}
source('dataGenerato.R')
#This generates data for species name, limb width, limb length, and observer name columns
```

```{r}
source('volumeEstimato.R')
#This calculates the volume of limbs using the formula for the volume of a cylinder
```

```{r}
#load in libraies 
library(dplyr)
library(ggplot2)
```

## Sort data

```{r}
sorted_data <- measurements %>%
  arrange(Limb_Volume_cm3) %>%
  arrange(Observer) %>%
  arrange(Species_Name)

#data is now sorted alphabetically by species, then observer, then in ascending order of limb volume
```

## Average Limb Volume by Species

```{r}
#get average limb volume for each species
average_volume <- measurements %>%
  group_by(Species_Name) %>%
  summarize(Average_Volume = mean(Limb_Volume_cm3))

#print results
print(average_volume)
```

## Number of Observations by Species and Observer

```{r}
#get number of observations for each combination of species/observer
counts <- measurements %>%
  group_by(Species_Name, Observer) %>%
  summarize (Count = n())

#create matrix for count data where x is observers and y is species
count_matrix <- matrix(counts$Count, nrow = 5, byrow = TRUE)
colnames(count_matrix)<- observer_vec
rownames(count_matrix)<- species_vec

#convert to table
count_table<- as.table(count_matrix)

#print result
print(count_table)
```

## Boxplot

```{r}
#create box plot of limb volume by species
boxplot<- ggplot(measurements, aes(x=Species_Name, y=Limb_Volume_cm3)) +
  geom_boxplot() +
  labs(title="Limb volume by species", x="Species", y="Limb Volume (cm3)")

print(boxplot)
```

**Figure 1.** Graph depicting limb volume data in cm3 for species Ally, Bally, Cally, Dally, and Elly. Volume was caclulated using the formula for the volume of a cylinder.

## Frequency Histograms

```{r}
#create faceted histogram of limb volume by species
histogram<- ggplot(measurements, aes(x=Limb_Volume_cm3)) +
  geom_histogram(bins = 5) +
  facet_wrap(~Species_Name)+
  labs(title="Limb volume frequency distribution by species", x="Limb Volume (cm3)", y="Count")

print(histogram)
```

**Figure 2** Graph depicting frequency distribution of limb volume data in cm3 for species Ally, Bally, Cally, Dally, and Elly. Volume was calculated using the formula for the volume of a cylinder.
